---
title: "HD-OCT"
author: "Pavel Dusek"
date: "19 11 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lmerTest)
library(ggplot2)
library(dplyr)
library(effectsize)
library(knitr)
```

## Software Info

Built with R version `r getRversion()`, lmerTest version `r packageVersion("lmerTest")`, effectsize version `r packageVersion("effectsize")`, ggplot2 version `r packageVersion("ggplot2")`, dplyr version `r packageVersion("dplyr")`.

## Data Description
We load the dataset:
```{r}
df <- read.csv('clean_data.csv')
colnames(df)
dim(df)
hd <- df %>% filter(group == 'patients') #patient group
hd_macular_nna <- hd %>% filter( !is.na(oct_macular_vol) ) #patient group without missing data in macular volume

```

To get descriptive statistics of the visits, we get parameters that are unique for each visits and remove duplicates:
```{r}
visits <- df[ , c("record_id", "age", "gender", "group", "uhdrs_tms", "uhdrs_tfc", "disease_length")]
visits <- visits[ !duplicated(visits), ]
```

To get descriptive statistics of the subjects, we get parameters that are unique for each subject and remove duplicates:
```{r}
subjects <- df[ , c("record_id", "gender", "group", "triplet_longer")]
subjects <- subjects[ !duplicated(subjects), ]
```

### How many eyes, visits and subjects are there?
How many measurements for each group?
```{r, echo=FALSE}
df %>% count( group )
```

### Do the groups differ in age?
Wilcox test:
```{r, echo=FALSE}
wilcox.test( age ~ group, visits)
```

### Do the groups differ in gender?
```{r, echo=FALSE}
fisher.test(table(subjects[ , c("gender", "group") ]))
```

### Other stats
```{r}
visits   %>% group_by( group ) %>% summarize( age_mean = mean(age), age_sd = sd(age) )
subjects %>% count( group )
visits   %>% count( group )
subjects %>% filter( group == 'patients' ) %>% count( gender )
subjects %>% filter( group == 'controls' ) %>% count( gender )
subjects %>% filter( group == 'patients' ) %>% summarize( cag = mean(triplet_longer), cag_sd = sd(triplet_longer) )
visits   %>% filter( group == 'patients' ) %>%
	summarize(
	    tfc = mean(uhdrs_tfc),
	    tfc_sd = sd(uhdrs_tfc),
	    tms = mean(uhdrs_tms),
	    tms_sd = sd(uhdrs_tms),
	    duration = mean(disease_length),
	    duration_sd = sd(disease_length)
	) 

```

## Group comparison mixed effects model

### RNFL G
```{r}
model1 <- lmer(oct_rnfl_g ~ group + age + gender + (1 | record_id), df )
summary(model1)
```

### RNFL T
```{r}
model2 <- lmer(oct_rnfl_t ~ group + age + gender + (1 | record_id), df )
summary(model2)
```

### Macular Volume
```{r}
model3 <- lmer(oct_macular_vol ~ group + age + gender + (1 | record_id), df )
summary(model3)
```

## HD group mixed effects model
We analyzed effect of UHDRS TMS and disease duration on RNFL G, RNFL T and macular volume in the HD group.
### RNFL G
```{r}
model4 <- lmer(oct_rnfl_g ~ uhdrs_tms + disease_length + (1 | record_id), hd )
summary(model4)
```

### RNFL T
```{r}
model5 <- lmer(oct_rnfl_t ~ uhdrs_tms + disease_length + (1 | record_id), hd )
summary(model5)
```

### Macular Volume
```{r}
model6 <- lmer(oct_macular_vol ~ uhdrs_tms + disease_length + (1 | record_id), hd )
summary(model6)
```
## P values false discovery rate
```{r echo=FALSE}
interpret <- function(p) {
	if (p < 0.001) {
		return("***")
	} else if (p < 0.01) {
		return("**")
	} else if (p < 0.05) {
		return("*")
	} else if (p < 0.10) {
		return(".")
	} else {
		return("")
	}
}
```
```{r}
p.values1 <- summary(model1)$coefficients[, "Pr(>|t|)"]
p.values2 <- summary(model2)$coefficients[, "Pr(>|t|)"]
p.values3 <- summary(model3)$coefficients[, "Pr(>|t|)"]
p.values4 <- summary(model4)$coefficients[, "Pr(>|t|)"]
p.values5 <- summary(model5)$coefficients[, "Pr(>|t|)"]
p.values6 <- summary(model6)$coefficients[, "Pr(>|t|)"]

p.values <- tibble(
		   name = c(
			    'g_intercept', 'g_group', 'g_age', 'g_gender',
			    't_intercept', 't_group', 't_age', 't_gender',
			    'm_intercept', 'm_group', 'm_age', 'm_gender',
			    'hdg_intercept', 'g_tms', 'g_disease_length',
			    'hdt_intercept', 't_tms', 't_disease_length',
			    'hdm_intercept', 'm_tms', 'm_disease_length'
			    ),
		   p = c(p.values1, p.values2, p.values3, p.values4, p.values5, p.values6)
	)
p.values <- p.values %>% mutate( p.BH = p.adjust(p, method="BH"), interpret = lapply(p.BH, interpret) )
kable(p.values, caption = 'Adjusted p values for false discovery rate.')
```

## Effect size
```{r}
g1 <- tibble(hedges_g(oct_rnfl_g      ~ group, data = df))
g2 <- tibble(hedges_g(oct_rnfl_t      ~ group, data = df))
g3 <- tibble(hedges_g(oct_macular_vol ~ group, data = df))
g <- rbind(g1, g2, g3)
g <- g %>% mutate( name = c( 'rnfl_g', 'rnfl_t', 'macular') )
kable(g, caption = 'Effect size of group on OCT parameters.')
```

## Macular Volume Post Hoc Analysis
```{r}
```

```{r}
model7 <- lmer(oct_macular_vol ~ group + age + gender + group*age + (1 | record_id), df )
summary(model7)
```

```{r}
model8 <- lmer( oct_macular_vol ~ triplet_longer * age + (1 | record_id), data = hd_macular_nna )
summary(model8)
```

Scatter plot of macular volume vs. age in patient and control group.
```{r}
ggplot( data = df, aes(x = age, y = oct_macular_vol)) + geom_point(aes(color=group)) + geom_smooth(aes(color=group), method='lm')
```

Correlation analysis of macular volume and CAG triplet extension in patient group.
```{r}
cor.test(hd_macular_nna$oct_macular_vol, hd_macular_nna$triplet_longer)
```

Correlation analysis of macular volume and age in patient group.
```{r}
cor.test(hd_macular_nna$oct_macular_vol, hd_macular_nna$age)
```
